### 2014-09-24 Studies/2011_San_Francisco/Report/San_Francisco_Fire.tex
San Francisco: Edits to fire and HRR section

To estimate the type of fuel, fire size, and structure ventilation, on-scene photographs and post-incident reports were used. Based on these sources of information, the FDS model includes three source fires: an initial couch fire in the basement, a secondary flashover fire that involved additional furniture items in the basement, and a third fire that occurred on the exterior basement balcony. The basement primarily contained upholstered polyurethane furniture fuel items, and the balcony was constructed of wood. The HRR of the initial couch fire is based on an experimentally measured HRR for a mattress fire~\cite{madrzykowski2009fire} to represent the fire growth rate of a similar household furniture item. The primary fuel packages were located in the basement as shown in Fig.~\ref{fig:fuel_placement}. In this figure, the numbered labels indicate the order in which the fuel items were burning: 1) the initial couch fire, 2) the secondary furniture fuel items, and 3) the balcony fire.

@@

@@

\noindent The HRR for the various fuel packages was prescribed in FDS as follows.

\begin{enumerate}
\item For the inital couch fire in the basement, the HRR from 0~s to 327~s followed the experimentally measured mattress HRR. At 300~s, the HRR was approximately 3~MW. From 327~s until 540~s (the end of the simulation), the HRR was a constant 5~MW.

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifying assumptions were made. The first assumption is that the initial couch fire in the basement was specified to start at the beginning of the simulation. The second assumption is that the sources of the gas phase polyurethane and wood fuel have constant areas at fixed locations (five furniture items in the basement interior for the polyurethane fuel, and the exterior basment balcony for the wood fuel). The following lines define the source fires in the FDS input file:

&SURF ID='WOOD_EXTERIOR',
    SPEC_ID='WOOD',
    MASS_FLUX = 0.003049,
    RAMP_MF='WOOD_EXTERIOR_RAMP' /
\end{lstlisting}
Note that the mass flux values here are unique for each fuel item and are specific to the input files used for these simulations. The key point is that each fuel item was modeled as a 3D object with flat surfaces from which a specified amount of fuel flows, mixes with air, and burns.

@@

In the FDS simulations, a prescribed HRR is specified, and FDS determines the amount of combustion that occurs throughout the simulation based on the amount of available fuel and oxygen at a given location. As a result, the actual HRR in FDS can be different than the prescribed HRR based on the ventilation conditions.

Figure~\ref{fig:hrr_w_fds} shows the HRR vs. time based on the fires that were described in Section~\ref{sec:fires}. The vertical dashed line represents the time at which the first basement window on the rear side of the structure failed (342~s). The maximum size of the fire used in this study was selected such that the conditions would become ventilation limited to minimize computational time while aligning with the timeline observations from the fire scenario. Note that the rear window failures were manually specified in accordance with the event timeline rather than failure temperature criteria.

@@

In Fig.~\ref{fig:hrr_w_fds}, the solid line represents the HRR that was prescribed in the FDS simulation, and the dashed line represents the actual HRR that was predicted by FDS based on the ventilation conditions. From 0~s to 300~s, the predicted HRR is in agreement with the prescribed HRR because there is an adequate amount of oxygen for all of the fuel to combust. From 300~s to 342~s (the time at which the first basement window failed), the predicted HRR is lower than the prescribed HRR because the fire in the structure becomes ventilation limited. The maximum predicted HRR value is approximately 7~MW before the fire becomes ventilation limited and begins to decay. The HRR decreased to approximately 2~MW as the amount of oxygen inside the structure becomes limited. At this time, there is an insufficient amount of oxygen in the structure to burn all of the fuel; therefore, localized burning occurs near the ventilation openings where ambient air is available.

At a time of 342~s, the first window on the rear side of the structure fails, and additional windows fail within the next two minutes. As additional window failures occur (see Table~\ref{tab:ventilation_timeline}), the abrupt change in ventilation causes the HRR to increase to approximately 32~MW before it reaches a steady-state value of approximately 20~MW. The predicted HRR increases above the prescribed HRR value because the unburned fuel and smoke that accumulated in the structure during ventilation-limited conditions is now mixing with ambient air from the failed window openings and burning. Figure~\ref{fig:timeline_comparison} shows a time series comparison of on-scene photographs and Smokeview snapshots of flaming combustion that occurs through the failed windows on the rear side of the structure.

\begin{figure}[!ht]
\includegraphics[height=8.25in]{../Figures/Timeline_Comparison}
\caption{Comparison of on-scene photographs and Smokeview snapshots of exterior combustion occurring on the rear side of the structure.}
\label{fig:timeline_comparison}
\end{figure}

>>>>

To estimate the type of fuel, fire size, and structure ventilation, on-scene photographs and post-incident reports were used. Based on these sources of information, the FDS model includes three source fires: 1) an initial couch fire in the basement, 2) a flashover fire that involved additional furniture items in the basement, and 3) a fire that occurred on the exterior basement balcony. The basement primarily contained upholstered polyurethane furniture fuel items, and the balcony was constructed of wood. The HRR of the initial couch fire is based on an experimentally measured HRR for a mattress fire~\cite{madrzykowski2009fire} to represent the fire growth rate of a similar household furniture item.

The primary fuel packages were located in the basement as shown in Fig.~\ref{fig:fuel_placement}. In this figure, the numbered labels indicate the order in which the fuel items were burning: 1) the initial couch fire, 2) the secondary furniture fuel items, and 3) the balcony fire.

@@

Since the time at which the fire started is not known relative to the notification of a fire or how fast the fire spread, several simplifying assumptions were made. One assumption is that the initial couch fire in the basement was specified to start at the beginning of the simulation. The beginning of the simulation ($t=0$) was selected to correspond to an on-scene time of 10:53:00 (see Table~\ref{tab:incident_timeline}). The HRRs for the three primary source fires was prescribed in FDS as follows.

\begin{enumerate}
\item For the initial couch fire in the basement, the HRR from 0~s to 327~s followed the experimentally measured mattress HRR. At 300~s, the HRR was approximately 3~MW. From 327~s until 540~s (the end of the simulation), the HRR was a constant 5~MW.

Another simplification of the model is that the sources of the gas phase polyurethane and wood fuel have constant areas at fixed locations (five furniture items in the basement interior for the polyurethane fuel, and the exterior basement balcony for the wood fuel). The following lines define the source fires in the FDS input file:

&SURF ID='WOOD_EXTERIOR',
    SPEC_ID='WOOD',
    MASS_FLUX = 0.003049,
    RAMP_MF='WOOD_EXTERIOR_RAMP' /
\end{lstlisting}

Note that the mass flux values here are unique for each fuel item and are specific to the input files used for these simulations. The key point is that each fuel item was modeled as a 3D object with flat surfaces from which a specified amount of fuel flows, mixes with air, and burns.

@@

@@

In an FDS simulation, a prescribed HRR is specified, which results in a specified amount of fuel vapor (or pyrolyzate) being released from a fuel surface. FDS then determines the amount of combustion that occurs throughout the simulation based on the amount of available fuel and oxygen at a given location. As a result, the actual HRR in FDS can be different than the prescribed HRR based on the ventilation conditions.

Figure~\ref{fig:hrr_w_fds} shows the HRR vs. time based on the fires that were described in Section~\ref{sec:fires}. The vertical dashed line represents the time at which the first basement window on the rear side of the structure failed (342~s). Additional windows on the rear side of the basement continued to fail within the next two minutes. The maximum size of the fire used in this study was selected such that the conditions would become ventilation limited to minimize computational time while aligning with observations from the actual fire scenario. Note that the rear window failures were manually specified in accordance with the event timeline rather than failure temperature criteria.

@@

In Fig.~\ref{fig:hrr_w_fds}, the solid line represents the HRR that was prescribed in the FDS simulation, and the dashed line represents the actual HRR that was predicted by FDS based on the ventilation conditions. From 0~s to 300~s, the predicted HRR was in agreement with the prescribed HRR because there was an adequate amount of oxygen for all of the fuel to combust. From 300~s to 342~s (the time at which the first basement window failed), the predicted HRR was lower than the prescribed HRR because the fire conditions in the structure became ventilation limited. The maximum predicted HRR was approximately 7~MW before the fire conditions became ventilation limited and started to decay. The HRR decreased to approximately 2~MW as the amount of oxygen inside the structure became limited. At this time, there was an insufficient amount of oxygen in the structure to burn all of the fuel; therefore, localized burning occured near the ventilation openings where ambient air was available.

At a time of 342~s, the first window on the rear side of the structure failed, and additional windows failed within the next two minutes. As additional window failures occurred (see Table~\ref{tab:ventilation_timeline}), the abrupt change in ventilation caused the HRR to increase to approximately 32~MW before it reached a steady-state value of approximately 20~MW. The predicted HRR increased above the prescribed HRR because the unburned fuel and smoke that accumulated in the structure during ventilation-limited conditions was mixing with ambient air from the failed window openings and burning.

Figure~\ref{fig:timeline_comparison} shows a time series comparison of on-scene photographs and Smokeview snapshots of flaming combustion that occured through the failed windows on the rear side of the structure. In this figure, the first photograph is shown at a simulation time of 321~s (approximately 20~s before the first rear basement window failed), the second photograph is shown at a simulation time of 349~s (approximately 10~s after the first rear basement window failed), the third photograph is shown at a simulation time of 440~s (approximately when the third rear basement window failed), and the fourth photograph is shown at a simulation time of 450~s (approximately when all of the rear basement windows failed).

\begin{figure}[!ht]
\includegraphics[height=8.0in]{../Figures/Timeline_Comparison}
\caption{Comparison of on-scene photographs and Smokeview snapshots of exterior combustion occurring on the rear side of the structure. The on-scene time is indicated on the left photograph, and the corresponding simulation time is indicated on the right Smokeview snapshot. For more information on the fire incident and simulation timeline, refer to Table~\ref{tab:combined_timeline}.}
\label{fig:timeline_comparison}
\end{figure}
